### 2025-04-23 glossary.tex

% glossary.tex
\newglossaryentry{FFT}{
    name={FFT},
    description={Fast Fourier Transforms}
}
\newglossaryentry{STFT}{
    name={STFT},
    description={Short Time Fourier Transform, a Fourier-based technique used in audio signal processing}
}

\newglossaryentry{DSP}{
    name={DSP},
    description={Digital Signal Processing}
}

\newglossaryentry{MSS}{
name = {MSS},
description={Multi-Scale Spectrogram}
}

\newglossaryentry{HC}{
name = {HC},
description={Hill-Climber algorithm}
}
\newglossaryentry{JTFS}{
name = {JTFS},
description={Joint-time Frequency Scattering}
}

\newglossaryentry{SIMSE}{
name = {SIMSE},
description = {Structured Similarity Metric}
}

\newglossaryentry{DTW}{
name = {DTW},
description={Dynamic Time Warping}
}

\newglossaryentry{DDSP}{
    name={DDSP},
    description={Differentiable Digital Signal Processing, overlap of methods from signal processing with machine learning}
}

\newglossaryentry{FM}{
    name={FM},
    description={A method of sound synthesis where the frequency of one oscillator modulates the frequency of another}
}

\newglossaryentry{Gradient descent}{
    name={Gradient descent},
    description={An optimization algorithm used to minimize a function by iteratively moving towards the lowest point}
}

\newglossaryentry{GA}{
name = {GA},
description={genetic algorithm}
}

% \newglossary{HC}{
% name = {HC},
% description={Hill climber}

% }

\newglossaryentry{NN}{
name = {NN},
description = {Neural network}
}

\newglossaryentry{VST}{
name = {VST},
description = {Virtual Studio Technology. Virtual software synthesizers which can run standalone or in a DAW}
}

\newglossaryentry{DAW}{
name = {DAW},
description = {Digital Audio Workstation}
}

\newglossaryentry{CI}{
name = {CI},
description = {Confidence interval}
}

\newglossaryentry{LFO}{
name = {LFO},
description = {low frequency oscillator}
}

\newglossaryentry{MFCC}{
name = {MFCC},
description = {Mel-frequency cepstral coefficients}
}

\newglossaryentry{AM}{
name = {AM},
description = {Amplitude modulation synthesis. A method of sound synthesis where one oscillator modulates the amplitude of another
}
}

\newglossaryentry{ADSR}{
    name = {ADSR},
    description= {Attack, delay, sustain, release. An envelope that shapes a sound. }
}

\newglossaryentry{OSR}{
name = {OSR},
description = {Open set recognition}
}

\newglossaryentry{RL}{
name = {RL},
description = {Reinforcement learning}
}

\newglossaryentry{SGD}{
name = {SGD},
description = {Stochastic gradient descent}
}

\newglossaryentry{DL}{
name = {DL},
description = {Deep learning}
}

\newglossaryentry{SOTA}{
name = {SOTA},
description = {State of the art}
}

\newglossaryentry{ML}{
name = {ML},
description = {Machine learning}
}

\newglossaryentry{SNPK}{
name = {SNPK},
description = {Non-Parametric Scott-Knott}
}